---
title: "Food Sustanibility"
author: "Declan Oberlies"
date: '`r Sys.Date()`'
output: ioslides_presentation
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
# load packeges ----
library(ggplot2)
library(dplyr)
library(readr)
library(leaflet)
library(patchwork)
library(tidyr)
library(purrr)
library(knitr)
library(plotly)
`%not_in%` <- negate(`%in%`)
```

## Sustainable Development Goal

SDG 2: eliminate hunger

## Sustainable Development Goal

SDG 2: eliminate hunger

\- grow more food

## Sustainable Development Goal

SDG 2: eliminate hunger

\- grow more food

\- less people

## Sustainable Development Goal

SDG 2: eliminate hunger

\- grow more food

\- ~~less people~~

## Sustainable Development Goal

SDG 2: eliminate hunger

\- grow more food

\- ~~less people~~

\- use food more efficiency

\- avoid supply chain disruptions

## 

```{r, echo=FALSE}
url <- "https://ourworldindata.org/explorers/global-food.csv?v=1&csvType=full&useColumnShortNames=true&Food=All+food&Metric=Food+available+for+consumption&Per+capita=true"

calories <- read_csv(url) %>%
  rename(kilocalories_day_capita = total__00002901__food_available_for_consumption__0664pc__kilocalories_per_day_per_capita, 
         country = entity) %>%
  filter(country %not_in% c("World",
                           "North America",
                           "South America",
                           "Africa",
                           "Asia",
                           "Ociana",
                           "Europe"),
         !grepl("\\(FAO\\)$", country))
# View(calories)

calories %>% 
  filter(year == 2023) %>% 
  arrange(desc(kilocalories_day_capita)) %>%
  head(6) %>% 
  select(country, kilocalories_day_capita) %>%
  kable(col.names = c("Country", "kilocalories/day/capita"))
```

## 

```{r}
url <- "https://ourworldindata.org/grapher/population-with-un-projections.csv?v=1&csvType=full&useColumnShortNames=true"

population <- read_csv(url) %>%  
  rename(country = entity,
         population = population__sex_all__age_all__variant_estimates) %>%
  filter(country %not_in% c("World",
                           "North America",
                           "South America",
                           "Africa",
                           "Asia",
                           "Ociana",
                           "Europe"),
         !grepl("\\(FAO\\)$", country))
joincal <- left_join(calories, population, by = c('country', 'year'))

joincal %>% 
  filter(year == 2023) %>% 
  arrange(desc(kilocalories_day_capita)) %>%
  head(6) %>% 
  select(country, kilocalories_day_capita, population) %>%
  kable(col.names = c("Country", "kilocalories/day/capita", "Population"))

```

## 

```{r}
joincal %>%
  mutate(kilocalories_day = kilocalories_day_capita*population) %>%
  filter(year == 2023) %>% 
  arrange(desc(kilocalories_day_capita)) %>%
  head(6) %>% 
  select(country, kilocalories_day) %>%
  kable(col.names = c("Country", "kilocalories/day"))
```

## 

```{r}
joincal %>%
  mutate(kilocalories_day = kilocalories_day_capita*population) %>%
  filter(year == 2023, 
         country %in% c('United States', 'China', 'India')) %>%
  select(country, population, kilocalories_day_capita, kilocalories_day) %>%
  kable(col.names = c("Country", "Population", "kilocalories/day/capita", "kilocalories/day"))
```

## 

```{r}
# Land use  ====================================================================
# https://ourworldindata.org/land-use#how-has-global-land-use-changed-over-the-long-term
url <- 'https://ourworldindata.org/grapher/land-use-over-the-long-term.csv?v=1&csvType=full&useColumnShortNames=true'
Land_use <- read_csv(url)
land_use <-
  Land_use %>% 
  filter( year %in% 2023, 
          entity %in% c('United States','China','India')) %>% 
  rename(uopp = uopp_c,
         cropland = cropland_c,
         rangeland = grazing_c) %>%
   pivot_longer(cols = uopp:cropland,
                names_to = "land_type",
                values_to = "ha")

joincal2 <- joincal %>%
  mutate(kilocalories = kilocalories_day_capita*population*365) %>%
  rename(entity = country)
Land_use %>%
  filter(year %in% 2023, 
         entity %in% c('United States','China','India')) %>%
  mutate(total_ha = cropland_c + grazing_c) %>%
  inner_join(joincal2, by = c("entity", "year")) %>%  
  mutate(ratio = kilocalories/total_ha) %>%
  select(entity,ratio) %>%
  kable(col.names = c("Country", "kilocalories/ha"))
         
```

## 

```{r echo=FALSE}
ggplotly(
  ggplot(land_use,
       aes(x = entity,
           y = (ha/1000000),
           fill = land_type)) +
  geom_col() +
  labs(y = "area in millions of ha",
       x = "Country",
       title = "Land Use")
)
```

## 

```{r}
url <- "https://ourworldindata.org/grapher/land-area-per-crop-type.csv?v=1&csvType=full&useColumnShortNames=true"
Crop <- read_csv(url)
crop <- 
  Crop %>% 
  filter( year %in% 2023, 
          entity %in% c('United States','China','India')) %>%
  mutate(country = entity,
         maize = maize__00000056__area_harvested__005312__hectares,
         wheat = wheat__00000015__area_harvested__005312__hectares,
         potatoes = potatoes__00000116__area_harvested__005312__hectares,
         sugar = sugar_cane__00000156__area_harvested__005312__hectares,
         soybeans = soybeans__00000236__area_harvested__005312__hectares,
         beans = beans__dry__00000176__area_harvested__005312__hectares,
         rice = rice__00000027__area_harvested__005312__hectares,
         rapeseed = rapeseed__00000270__area_harvested__005312__hectares,
         other = sum(sugar_beet__00000157__area_harvested__005312__hectares,
                     cocoa_beans__00000661__area_harvested__005312__hectares,
                     cassava__00000125__area_harvested__005312__hectares,
                     tomatoes__00000388__area_harvested__005312__hectares,
                     peas__dry__00000187__area_harvested__005312__hectares,
                     bananas__00000486__area_harvested__005312__hectares,
                     oranges__00000490__area_harvested__005312__hectares,
                     apples__00000515__area_harvested__005312__hectares,
                     rye__00000071__area_harvested__005312__hectares,
                     barley__00000044__area_harvested__005312__hectares,
                     na.rm = TRUE)
         ) %>%
  pivot_longer(cols = maize:other,
               names_to = "Crop",
               values_to = "ha") %>%
  mutate(Crop = 
           factor(Crop,
                            levels = c("maize",
                                       "wheat",
                                       "potatoes",
                                       "sugar",
                                       "soybeans",
                                       "beans",
                                       "rice",
                                       "rapeseed", 
                                       "other"))
         )

ggplotly(
  ggplot(crop,
       aes(x = country,
           y = (ha/1000000),
           fill = Crop)
       ) +
  geom_col() +
    labs(x = "Country",
         y = "area in millions of ha",
         title = "Crop Area Use")
  )
```

## 

```{r}
url <- "https://ourworldindata.org/grapher/ghg-emissions-by-sector.csv?v=1&csvType=full&useColumnShortNames=true"

co2 <- read_csv(url) %>%
  filter(year == 2023,
         entity %in% c("United States",
                       "China",
                       "India")
                       ) %>%
  select(entity,
         year,
         agriculture_ghg_emissions) %>%
  rename(country = entity,
         agriculture_emissions = agriculture_ghg_emissions)

co2 %>%
  select(country,
         agriculture_emissions) %>%
  kable()

```

## Conclusion

**No one is perfect**

| USA | India |
|----|----|
| high caloric supply to a large population | supply a massive population with minimal land |
| BUT too much meat | good crop variability |
| BUT low crop variability | BUT high agricultural CO~2~ |

## **The End!**





